Combined effects of protein expression variance and correlation on multicomponent systems

### 2020-03-04 Analysis.Rmd
updated

totVarPair <- function(x){
  A <- as.character(x[1])
  B <- as.character(x[2])
  varA <- as.numeric(protVar[protVar$Uniprot == A, 2])
  varB <- as.numeric(protVar[protVar$Uniprot == B, 2])
  corAB <- cor[A,B]
  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
  return(tot)
}

>>>>

totVarPair <- function(x){
  A <- as.character(x[1])
  B <- as.character(x[2])
  varA <- as.numeric(protVar[protVar$Uniprot == A, 3])
  varB <- as.numeric(protVar[protVar$Uniprot == B, 3])
  corAB <- cor[A,B]
  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
  return(tot)
}

### 2020-03-04 Manuscript/Manuscript.Rmd
updated

  word_document: null
z
documentclass: article
capsize: normalsize
geometry: margin=1in
fontsize: 11pt
spacing: doublespacing
bibliography: 
  - references.bib
  - knitcitations.bib
csl: cell.csl
abstract: Protein expression variation leads to phenotypic variance between cells. This has been demonstrated in cell signaling and differentiation decisions. Additionally coordinated expression of proteins between cells can tune signaling pathways either towards a more binary or analog modality. Though there is some evidence in bulk cell measurements and in bacteria that certain heteromeric subunits or metabolic pathways may be expressed in a coordinated fashion (i.e. operons in bacteria), there has not been a direct measurement of coordinated expression of proteins (independent of TFs) in metazoans (vertebrates?). Here, we measure cell-to-cell variability of relative protein abundance using quantitative proteomics of individual *Xenopus laevis* eggs and show that proteins involved in metabolic pathways or members of heteromeric complexes tend to have high correlations with other members of those pathways/complexes. Our previous work highlighted the fact that correlated expression increases the total variation of a pathway, so one would reason that certain pathways or complexes would need to compensate of this extra source of by reducing the variation of expression of these pathways or complexes. To test this we computed total variance score that took into account both the coefficient of variance and correlations between proteins in a pathway and found that the lower 10% of GO terms were highly enriched for metabolic pathways. When we looked at the relationship between CV and R between these GO terms we found a negative relationship between them, demonstrating that increased correlation needs to come at the expense of decreased variance. Simple molecular models of heteromeric complexes and metabolic pathways demonstrate that this tradeoff can result in higher efficiencies in both function and reduced energy waste. Together, our study argues for a control principle whereby coordinated expression of proteins in a pathway can require lower variance in order to reduce the overall pathway variance, enabling accurate control of active complexes or metabolic pathway activity.
---

@@

library("knitr")

@@

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

@@

The coordinated expression of proteins is vital across cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues. The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases). These processes, as well as the molecules they target, are all subject to noise, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells. Variability and coordinated expression (correlation) of proteins leads to increased population level control in binary decisions, and a decreased ability to execute analog signaling.

@@

To fetch bibliographic metadata automatically from the web. For example, citing a paper can be as easy as providing its DOI `r citep("10.1016/j.tree.2006.03.016")` or even just a few keywords `r citep(c("Heterogeneous ribosomes preferentially translate distinct subpools of mRNAs genome-wide","Controlling low rates of cell differentiation through noise and ultrahigh feedback"))`.

@@

Activated Xenopus laevis eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins. Expression of these proteins were largely invariant across the cell cycle, revealing that these highly expressed genes are likely not regulated by cell cycle processes. Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S1).

To determine the variance across these proteins, we calculated the coefficient of variation (CV) using all time points collected. This showed a wide range of variation containing multiple distributions (Fig 1B), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2). In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…

@@

@@

$$var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$$

>>>>

  word_document: null
documentclass: article
capsize: normalsize
geometry: margin=1in
fontsize: 11pt
spacing: doublespacing
bibliography: 
  - references.bib
  - knitcitations.bib
csl: cell.csl
abstract: Protein expression variation leads to phenotypic variance between cells. This has been demonstrated in cell signaling and differentiation decisions. Additionally coordinated expression of proteins between cells can tune signaling pathways either towards a more binary or analog modality. Though there is some evidence in bulk cell measurements and in bacteria that certain heteromeric subunits or metabolic pathways may be expressed in a coordinated fashion (i.e. operons in bacteria), there has not been a direct measurement of coordinated expression of proteins (independent of TFs) in metazoans (vertebrates?). Here, we measure cell-to-cell variability of relative protein abundance using quantitative proteomics of individual *Xenopus laevis* eggs and show that proteins involved in metabolic pathways or members of heteromeric complexes tend to have high correlations with other members of those pathways/complexes. Our previous work highlighted the fact that correlated expression increases the total variation of a pathway, so one would reason that certain pathways or complexes would need to compensate of this extra source of by reducing the variation of expression of these pathways or complexes. To test this we computed total variance score that took into account both the coefficient of variance and correlations between proteins in a pathway and found that the lower 10% of GO terms were highly enriched for metabolic pathways. When we looked at the relationship between CV and R between these GO terms we found a negative relationship between them, demonstrating that increased correlation needs to come at the expense of decreased variance. Simple molecular models of heteromeric complexes and metabolic pathways demonstrate that this tradeoff can result in higher efficiencies in both function and reduced energy waste. Together, our study argues for a control principle whereby coordinated expression of proteins in a pathway can require lower variance in order to reduce the overall pathway variance, enabling accurate control of active complexes or metabolic pathway activity.
---

@@

library(knitr)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
library(cowplot)
library(patchwork)
source("../functions.R")
library(knitcitations)

cleanbib()   
cite_options(citation_format = "pandoc")

@@

```{r data_import, include=FALSE}

# Import Data
files = list.files('../TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('pepRaw')){
    rm(pepRaw)
  } 
  if(!exists('pepRaw')){
    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(pepRaw)
    pepRaw = rbind(pepRaw, temp)
    rm(temp)
  }
}
pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                  `Quan Usage` == 'Use',
                                                                  Confidence != 'Low')

```{r data_organization, include=FALSE}
pepTidy <- pepRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
  group_by(file, `Annotated Sequence`) %>% 
  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
  dplyr::select(file, run, pepID, everything()) %>%
  gather(key = label, value = intensity, '126':'131') %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
```

```{r normalization, include=FALSE}
protImpute <- pepTidy %>% 
  group_by(peptide,run) %>% 
  mutate(intensity = intensity / intensity[label == "126"]) %>%
  ungroup() %>%
  filter(run != "ctrl", condition != "ctrl") %>%
  dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
  group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
  group_by(Uniprot) %>%
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  group_by(cell) %>% 
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance) %>%
  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

```{r summary_stats, include=FALSE}
protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cov) = protImpute$Uniprot
rownames(cov) = protImpute$Uniprot
```

@@

The coordinated expression of proteins is vital for many aspects of cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues. The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases). These processes, as well as the molecules they target, are all subject to noise, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells. Variability and coordinated expression (correlation) of proteins leads to increased population level control in binary decisions, and a decreased ability to execute analog signaling.

@@

@@

The specific requirments for coordinated protein expression of pathways or complexes can vary, with some requiring strict stoichiometric regulation with others having much more relaxed requirements. Strict stoichiometry scenerios require coordinated expression of proteins (Figure 1A, top), and others can have uncoordinated expression (Figure 1A bottom). At the single cell level, coordinated expression will result in a high correlation coefficient, whereas un-coordinated expression will result in a low correlation coefficient (Figure 1B). Even when the expression variation (standard deviation or coefficient of variation) and abundance is identical between these two scenerios, the total variance of a pathway or complex can be significantly different due to the variance sum law: $var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$ (Figure 1C). Our previous work has shown that this effect is important in single cell pathway activation dynamics `r citep("doi.org/10.15252/msb.20177997")`.

In order to study the relationship between protein expression variance and coexpression (correlated expression) of cellular pathways and complexes at the proteome level, we utilized *Xenopus laevis* eggs as a single cell model. Activated *Xenopus laevis* eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins. Expression of these proteins were largely invariant across the cell cycle, revealing that these highly expressed genes are likely not regulated by cell cycle processes. Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S1).

To determine the variance across these proteins, we calculated the variance and coefficient of variation (CV) using all time points collected. This showed a wide range of variation containing multiple distributions (Fig 1D), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2). In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…

@@

```{r figure_1, echo=FALSE, fig.height=6.5, fig.width=6.5, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}

set.seed(41)

low_cor <- corProts(2, 100, 0, 100, 30)
low_cor <- low_cor %>% as_tibble() %>% mutate(cor = "low_cor", time = 1:nrow(low_cor)) %>%
  rename(ProtA = "V1", ProtB = "V2")

high_cor <- corProts(2, 100, 0.8, 100, 30)
high_cor <- high_cor %>% as_tibble() %>% mutate(cor = "high_cor", time = 1:nrow(high_cor)) %>%
  rename(ProtA = "V1", ProtB = "V2")

p1 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
  ggplot(., aes(x = time, y = abundance, color = protein)) + geom_path() +
  facet_wrap(~cor, ncol = 1) +
  theme_minimal() +
  theme(legend.position = "bottom")

p2 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
  ggplot(., aes(x = protein, y = abundance, color = protein)) + 
  geom_boxplot() + facet_wrap(~cor, ncol = 1) +
  theme_minimal() +
  theme(legend.position = "bottom")

p3 <- rbind(low_cor, high_cor) %>%
  ggplot(., aes(x = ProtA, y = ProtB, color = cor)) + geom_point() +
  stat_ellipse() +
  theme_minimal() +
  theme(legend.position = "bottom")

p4 <- tibble(cor = c("High Cor","Low Cor"),
       var_sum = c(sqrt(30^2 + 30^2 + 2*0.8*30^2),
                   sqrt(30^2 + 30^2 + 2*0*30^2))) %>%
  ggplot(., aes(x = cor, y = var_sum, fill = cor)) + geom_bar(stat = "identity") +
  theme_minimal()

# pp1 <- p1 + p2 + plot_layout(widths = c(3,1))
# pp2 <- p3 + p4
# 
# pp1 / pp2 + plot_annotation(tag_levels = "A") & theme(legend.position = "none")

p1 / (p3 + p4) + plot_annotation(tag_levels = "A") & theme(legend.position = "none")

fig_legend_1 <- c("A) Simulated expression of two proteins over time in a single cell with a mean of 100 and a standard deviation of 30 with expression correlations of 0 and 0.8. B) Pairwise plots of the two proteins in the single cells between uncorrelated and correlated scenerios. C) The total variance between the uncorrelated and correlated scenerios from A and B.")

@@

### 2020-03-04 Manuscript/Manuscript.tex
updated
update

\documentclass[11pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Combined effects of protein expression variance and correlation on multicomponent systems},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

To fetch bibliographic metadata automatically from the web. For example,
citing a paper can be as easy as providing its DOI (Clark and Gelfand,
2006) or even just a few keywords (Ahrends et al., 2014; Shi et al.,
2017).

@@

Activated Xenopus laevis eggs were collected at 5 time points across the
first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each
time point. Using TMT multiplexing and mass spectrometry, we were able
to determine the relative abundance of more than 1300 proteins.
Expression of these proteins were largely invariant across the cell
cycle, revealing that these highly expressed genes are likely not
regulated by cell cycle processes. Additionally, a PCA analysis of these
eggs showed no discernible clustering on cell cycle time (Fig S1).

To determine the variance across these proteins, we calculated the
coefficient of variation (CV) using all time points collected. This
showed a wide range of variation containing multiple distributions (Fig
1B), many of which are consistent with our previous study of variation
using targeted mass spectrometry (Fig S2). In order to see if variation
was a regulated process, we grouped the protein variation by gene
ontology terms (GO terms) and plotted them in ranked order. We saw that
in general, processes xxx were\ldots{}

@@

@@

Variance Sum Law --- Independent Case

\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2\]

@@

\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2Cov(x,y)\]
\[\rho = \dfrac{Cov(x,y)}{\sigma_{_{x}} \sigma_{_{y}}}\]
\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2\rho \sigma_{_{x}} \sigma_{_{y}}\]

\[\sigma_{{}_{total}} = \sqrt{\sum\limits_{i=1}^{n} \; \sigma_{i}^{2} \; \sum\limits_{j\neq i} \; 2\rho_{_{i,j}}\sigma_{_{i}}\sigma_{_{j}}}\]

\[\varsigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}\]

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-ahrends2014controlling}{}%
Ahrends, R., Ota, A., Kovary, K.M., Kudo, T., Park, B.O., and Teruel,
M.N. (2014). Controlling low rates of cell differentiation through noise
and ultrahigh feedback. Science \emph{344}, 1384--1389.

\leavevmode\hypertarget{ref-Clark_2006}{}%
Clark, J.S., and Gelfand, A.E. (2006). A future for models and data in
environmental science. Trends in Ecology \& Evolution \emph{21},
375--380.

\leavevmode\hypertarget{ref-shi2017heterogeneous}{}%
Shi, Z., Fujii, K., Kovary, K.M., Genuth, N.R., Röst, H.L., Teruel,
M.N., and Barna, M. (2017). Heterogeneous ribosomes preferentially
translate distinct subpools of mRNAs genome-wide. Molecular Cell
\emph{67}, 71--83.e7.

>>>>

\documentclass[11pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Combined effects of protein expression variance and correlation on multicomponent systems},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

@@

The specific requirments for coordinated protein expression of pathways
or complexes can vary, with some requiring strict stoichiometric
regulation with others having much more relaxed requirements. Strict
stoichiometry scenerios require coordinated expression of proteins
(Figure 1A, top), and others can have uncoordinated expression (Figure
1A bottom). At the single cell level, coordinated expression will result
in a high correlation coefficient, whereas un-coordinated expression
will result in a low correlation coefficient (Figure 1B). Even when the
expression variation (standard deviation or coefficient of variation)
and abundance is identical between these two scenerios, the total
variance of a pathway or complex can be significantly different due to
the variance sum law:
\(var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})\)
(Figure 1C). Our previous work has shown that this effect is important
in single cell pathway activation dynamics (Kovary et al., 2018).

In order to study the relationship between protein expression variance
and coexpression (correlated expression) of cellular pathways and
complexes at the proteome level, we utilized \emph{Xenopus laevis} eggs
as a single cell model. Activated \emph{Xenopus laevis} eggs were
collected at 5 time points across the first cell cycle (0, 20, 40, 60,
and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing
and mass spectrometry, we were able to determine the relative abundance
of more than 1300 proteins. Expression of these proteins were largely
invariant across the cell cycle, revealing that these highly expressed
genes are likely not regulated by cell cycle processes. Additionally, a
PCA analysis of these eggs showed no discernible clustering on cell
cycle time (Fig S1).

To determine the variance across these proteins, we calculated the
variance and coefficient of variation (CV) using all time points
collected. This showed a wide range of variation containing multiple
distributions (Fig 1D), many of which are consistent with our previous
study of variation using targeted mass spectrometry (Fig S2). In order
to see if variation was a regulated process, we grouped the protein
variation by gene ontology terms (GO terms) and plotted them in ranked
order. We saw that in general, processes xxx were\ldots{}

@@

\begin{figure}
\centering
\includegraphics{output/figures/figure_1-1.pdf}
\caption{A) Simulated expression of two proteins over time in a single
cell with a mean of 100 and a standard deviation of 30 with expression
correlations of 0 and 0.8. B) Pairwise plots of the two proteins in the
single cells between uncorrelated and correlated scenerios. C) The total
variance between the uncorrelated and correlated scenerios from A and
B.}
\end{figure}

@@

@@

The sum of variance for a group of proteins \(X_{1},\dotsc,X_{n}\) with
dependent relationships:

\[\sigma_{{}_{total}} = \sqrt{var\left(\sum_{i=1}^n X_{i}\right)}\]

\[\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}\]

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Kovary_2018}{}%
Kovary, K.M., Taylor, B., Zhao, M.L., and Teruel, M.N. (2018).
Expression variation and covariation impair analog and enable binary
signaling control. Molecular Systems Biology \emph{14}.
